Mining Interaction Log Data in a Creative Arts Practice using Transition Matrices

### 2015-09-23 flux-entropy-paper/flux-entropy-paper.tex
adjusted comments to be in comments - address these tonight

In our corpus of iPad musical performances, logs of low-level touch-data record each performer's micro-gestures. 
Because it has been
established that musical performances on touch-screens can be chuncked into
vocabularies of continuous macro gestures that unfold over
several seconds of touch interaction~\cite{Martin:2014cr}, these logs
 can be thought of as a sequence of such
states~\cite{Pressing:1988uo} for each performer.
If a performers' touch interactions can be classified as belonging
to one of $m$ distinct states, then, over the course of a session, each
participant $X$ generates a length $N$ sequence of these interaction
states $X_n$ ($n = 1, \ldots, N$) which can be thought of as Markov
process.
If there is an (automated) classification process for identifying each
participant's state sequence from their interaction data, then we can
estimate the $m \times m$ transition matrix (TM), $P$, which characterises
this process by setting each entry, $p_{ij}$, to be the number of times
state $j$ follows state $i$ in the sequence. The transition activity
of the whole ensemble can be calculated by averaging the TM for each
performer. The matrix can be normalised so that the sum of all
elements is equal to one.

@@

For entropy, significant main effects were found due to session
context ($F(2,86) = 12.10, p < 0.001$), and instrumentation
($F(1,86) = 25.86$) (WHAT IS THE P VALUE HERE?). There was no main effect for session type although
 the interaction of session type and
instrumentation was found to be significant
($F(1,86) = 9.93, p<0.01$).

>>>>

In our corpus of iPad musical performances, logs of low-level touch-data record each performer's micro-gestures. 
Because it has been
established that musical performances on touch-screens can be chuncked into
vocabularies of continuous macro gestures that unfold over
several seconds of touch interaction~\cite{Martin:2014cr}, these logs
 can be thought of as a sequence of such
states~\cite{Pressing:1988uo} for each performer.
% TODO need to refine the meaning of this sentence - now Pressing
% doesn't make sense as a citation.
If a performers' touch interactions can be classified as belonging
to one of $m$ distinct states, then, over the course of a session, each
participant $X$ generates a length $N$ sequence of these interaction
states $X_n$ ($n = 1, \ldots, N$) which can be thought of as Markov
process.
If there is an (automated) classification process for identifying each
participant's state sequence from their interaction data, then we can
estimate the $m \times m$ transition matrix (TM), $P$, which characterises
this process by setting each entry, $p_{ij}$, to be the number of times
state $j$ follows state $i$ in the sequence. The transition activity
of the whole ensemble can be calculated by averaging the TM for each
performer. The matrix can be normalised so that the sum of all
elements is equal to one.

@@

For entropy, significant main effects were found due to session
context ($F(2,86) = 12.10, p < 0.001$), and instrumentation
($F(1,86) = 25.86$).
% TODO (WHAT IS THE P VALUE HERE?). 
There was no main effect for session type although
 the interaction of session type and
instrumentation was found to be significant
($F(1,86) = 9.93, p<0.01$).
